---
title: "DHCE Project"
author: "Tanuj 'Rick' Guha"
date: "`r Sys.Date()`"
output: powerpoint_presentation
---

## Introduction. 

In this presentation we will compare different standards to qualify student promotion to the 4th grade. The first standard is called the **Basic Criteria**, and the second, the **Expanded Criteria**.

## Criteria Definitions

### Here are defitions of the two criteria of student promotion:

* Basic Criteria: to be promoted to 4th grade is that a student must obtain a proficiency level of 3 or 4 on the Spring State Test in 3rd grade ELA.

* Expanded Criteria: to be promoted to 4th grade a student a student must meet either the basic criteria, or they score at least at the 40th percentile on the 3rd grade Spring ELA screener.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#loading libraries
library(tidyverse)
library(readxl)
library(openxlsx)
library(datadictionary)
library(sqldf)
library(VennDiagram)
library(ggplot2)
library(gridExtra)
library(grid)
library(ggpubr)

#loading the data
grades = read.xlsx("ResearchAnalyst_InterviewProject_2023.xlsx",  sheet = 'Data')

#extracting only ela grades 
ela = grades[grades$Subject == "ELA",]
```

```{r, include=FALSE}
#creating a dataset of all student that pass according to the basic criteria. 
basic = ela[ela$Test == 'State Test' & ela$Testing.Window == "SPRING" & ela$Proficiency.Level > 2, ]

#creating a dataset of all student score at least at the 40th percentile on the 3rd grade Spring ELA screener. 
expandedOnly = ela[ela$Test == 'Screener' & ela$Testing.Window == "SPRING" & ela$Percentile > 39, ]

#finally creating a dataset of all students that pass according to the expanded criteria
expanded <- rbind(basic, expandedOnly)
```

```{r, include=FALSE}
#while the above datasets are created for analysis within R, let us create an spreadsheet with information of students who were promoted based on the respective standards. 

ela$Basic = ela$Test == 'State Test' & ela$Testing.Window == "SPRING" & ela$Proficiency.Level > 2

ela$Expanded = (ela$Test == 'State Test' & ela$Testing.Window == "SPRING" & ela$Proficiency.Level > 2) | (ela$Test == 'Screener' & ela$Testing.Window == "SPRING" & ela$Percentile > 39)

#quick internal check => output should be 0
print(nrow(ela[ela$Basic == TRUE & ela$Expanded == FALSE, ]))

#exporting the dataset out as a csv
write.csv(ela, paste(getwd(),"gradePromotion.csv", sep = "\\"), row.names=FALSE)

```

## Calculating Retention Rates

```{r, include=FALSE}
rateB = nrow(basic) / nrow(ela)
rateE = nrow(expanded) / nrow(ela)
```

```{r, echo = FALSE}
print("Retention Rate when using Basic Criteria:")
print(rateB)
```

```{r, echo = FALSE}
print("Retention Rate when using Expanded Criteria:")
print(rateE)
```
```{r, include = FALSE}

#creating a stacked bar graph of students who passed under basic criteria
basic <- basic %>% 
  mutate(Pass = "Basic")
basicDis = ggplot(data = basic) +
  geom_bar(aes(x = Pass, fill = factor(Economically.Disadvantaged)), position = "fill") + theme(legend.position="none")

#creating a stacked bar graph of students who passed under basic criteria
expanded <- expanded %>% 
  mutate(Pass = "Expanded")
expandedDis = ggplot(data = expanded) + geom_bar(aes(x = Pass, fill = factor(Economically.Disadvantaged)), position = "fill") + theme(legend.position="none")

forLegend = ggplot(data = expanded) + geom_bar(aes(x = Pass, fill = factor(Economically.Disadvantaged)), position = "fill")
legend = get_legend(forLegend)

plotM = grid.arrange(basicDis, expandedDis, ncol=2)
```

## Affect of Expanded Criteria on Disadvantaged Students

Economically disadvantaged students make up a greater proportion of students who passed using the expanded criteria, when compared to students who passed using the basic criteria.

```{r, echo = TRUE}
grid.arrange(plotM, legend, nrow = 2)
```

## Analysis of Test Performance

Given the expansion criteria, one could ostensibly expect all students who meet 3 or 4 proficiency in the state proficiency should also score above 40 percentile in the screener. 

Then, it might be interesting to see students who manage to pass in the state test, but fail in the screener. 

```{r, include=FALSE}

bne = as.data.frame(setdiff(basic$Student.ID, expandedOnly$Student.ID))
enb = as.data.frame(setdiff(expandedOnly$Student.ID, basic$Student.ID))
cross = nrow(basic)-nrow(bne)
```

```{r, echo = FALSE}
grid.newpage()
draw.pairwise.venn(area1=nrow(basic), area2=nrow(expandedOnly),cross.area=cross,
                   category=c("State Test 3/4","Screener 40%ile"),fill=c("Red","Yellow"))
```

## Final Thoughts

### Findings:
* The retention Rate when using Basic Criteria is 0.2014388
* The retention Rate when using Expanded Criteria is 0.4334532
* Expanding the criteria is good for economically disadvantaged students
* 16 People __passed__ in the state test, but __failed__ the screener. 
