---
title: "DHCE Project"
author: "Rick"
date: "`r Sys.Date()`"
output: powerpoint_presentation
---

## Introduction. 

In this presentation we will compare different standards to qualify student promotion to the 4th grade. The first standard is called the **Basic Criteria**, and the second, the **Expanded Criteria**.

### Here are defitions of the two criteria of student promotion:

* Basic Criteria: to be promoted to 4th grade is that a student must obtain a proficiency level of 3 or 4 on the Spring State Test in 3rd grade ELA.

* Exapanded Criteria: to be promoted to 4th grade a student a student must meet either the basic criteria, or they score at least at the 40th percentile on the 3rd grade Spring ELA screener.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#loading libraries
library(tidyverse)
library(readxl)
library(openxlsx)
library(datadictionary)
library(sqldf)
library("VennDiagram")

#loading the data
grades = read.xlsx("ResearchAnalyst_InterviewProject_2023.xlsx",  sheet = 'Data')

#extracting only ela grades 
ela = grades[grades$Subject == "ELA",]
```

```{r, include=FALSE}
#creating a dataset of all student that pass according to the basic criteria. 
basic = ela[ela$Test == 'State Test' & ela$Testing.Window == "SPRING" & ela$Proficiency.Level > 2, ]

#creating a dataset of all student score at least at the 40th percentile on the 3rd grade Spring ELA screener. 
expandedOnly = ela[ela$Test == 'Screener' & ela$Testing.Window == "SPRING" & ela$Percentile > 39, ]

#finally creating a dataset of all students that pass according to the expanded criteria
expanded <- rbind(basic, expandedOnly)
```

```{r, include=FALSE}
rateB = nrow(basic) / nrow(ela)
rateE = nrow(expandedOnly) / nrow(ela)
```

```{r}
```

## conitnue 3

```{r, include=FALSE}

bne = as.data.frame(setdiff(basic$Student.ID, expandedOnly$Student.ID))
enb = as.data.frame(setdiff(expandedOnly$Student.ID, basic$Student.ID))
cross = nrow(basic)-nrow(bne)
```

## conitnue 4

```{r}
print ("Is ")
print((nrow(basic)-nrow(bne)) == (nrow(expandedOnly)-nrow(enb)))

```
## pretty

```{r, include=FALSE}
grid.newpage()
draw.pairwise.venn(area1=nrow(basic), area2=nrow(expandedOnly),cross.area=cross,
                   category=c("Basic","expandedOnly"),fill=c("Red","Yellow"))
```


```{r, include=FALSE}
basicDR = nrow(basic[basic$Economically.Disadvantaged == 'Yes', ]) / nrow(basic)
expandedOnlyDR = nrow(expandedOnly[expandedOnly$Economically.Disadvantaged == 'Yes', ]) / nrow(expandedOnly)
```


Student in expandedOnly but not in basic
Student in basic, but not in expandedOnly
How many more economically disadvantaged students does expanding the criteria help? 
